---
title: "hw2"
author: "Chen Chen"
date: "March 28, 2018"
output: md_document
---


## step1: read the data file in to R
```{r}
data<- read.csv("activity.csv")
```

## step2: calculate the mean total number of steps taken per day
   mean total number=total number of steps/number of days
   
   Let's calculate the total number of steaps taken every day. It is stored in the data frame total.steps.daily
```{r}
# calculate the sum of steps by date
total.steps.daily<- sapply(split(data$steps,data$date),function(x){sum(x,na.rm = TRUE)})
total.steps.daily<-data.frame(toal.steps=total.steps.daily)
total.steps.daily[,2]<-data.frame(date=sapply(split(data$date,data$date),function(x){x[1]}))
total.steps.daily$date=as.Date(as.character(total.steps.daily$date),"%Y-%m-%d")
```

  Then a histogram about the total number of steps taken every day is shown
```{r}
hist(total.steps.daily$toal.steps,breaks = 10,xlab = "total steps every day",main = "histogram of total steps every day")
```

 The mean of the total number of steps every day is:
```{r}
 mean(total.steps.daily$toal.steps,na.rm = TRUE)
```
  
  The median of the total number of steps every day is
```{r}
median(total.steps.daily$toal.steps,na.rm = TRUE)
```


## step3: find the average daily activity pattern
  If we calculate the mean steps per interval, and plot the mean versus intervals, we are going to see this figure.
```{r}
interval.steps<-data.frame(average.step=sapply(split(data$steps,as.factor(data$interval)),function(x){mean(x,na.rm = TRUE)}))
interval.steps[,2]<-data.frame(interval=sapply(split(data$interval,as.factor(data$interval)),function(x){x[1]}))
with(interval.steps,plot(interval, average.step,type = "l",col="blue"))
```



  The interval with the maximal of steps is
```{r}
interval.steps$interval[which.max(interval.steps$average.step)]
```


## step4: imputing missing values
  The number of missing values in the dataset is:
```{r}
nrow(data)-sum(complete.cases(data))
```

  If we use the mean value of each interval to fill in the missing places, we would have the new dataset y:
```{r}
y<-data
y<-merge(y,interval.steps,by="interval")
index<-complete.cases(y)
y$steps[!index]<-y$average.step[!index]    ##fill in the new data
```

 If we redo the calculation in Step 1, the histogram will be,
```{r}
 total.steps.daily<- sapply(split(y$steps,y$date),function(x){sum(x,na.rm = TRUE)})
 total.steps.daily<-data.frame(toal.steps=total.steps.daily)
 total.steps.daily[,2]<-data.frame(date=sapply(split(y$date,y$date),function(x){x[1]}))
 total.steps.daily$date=as.Date(as.character(total.steps.daily$date),"%Y-%m-%d")

 hist(total.steps.daily$toal.steps,breaks = 10,xlab = "total steps every day",main = "histogram of total steps every day")
```
 

 The mean of the total number of steps every day is:
```{r}
 mean(total.steps.daily$toal.steps,na.rm = TRUE)
```
  
  The median of the total number of steps every day is
```{r}
median(total.steps.daily$toal.steps,na.rm = TRUE)
```

They are all different from the results in Step 1. The mean and the median are larger.


## Step 5: Are there differences in activity patterns between weekdays and weekends?

  Let's creat a new factor variable in data frame y to indicate whether that day is weekday or weekend.
```{r}
y$date<-as.Date(as.character(y$date))
index1<-is.element(weekdays(y$date),c("Saturday","Sunday"))
 y$week<-NA
y$week[index1]<-"weekends"
y$week[!index1]<-"weekdays"
y$week<-as.factor(y$week)
head(y)
```


  The patterns in Step3 are splited by whether it is on weekdays or weekends.

```{r}
z<-lapply(split(y,y$week),function(x){aggregate(x$steps,by=list(x$interval),FUN=mean)})
z[[1]]$week<-"weekdays"
z[[2]]$week<-"weekends"
average.steps.by<-rbind(z[[1]],z[[2]])
names(average.steps.by)<-c("interval","average.steps","week")
library("ggplot2")
qplot(data = average.steps.by,x=interval,y=average.steps,facets = .~week,geom = "line",color=week)
```


























